---
title: "Filters and joins"
author: "Max Kagan"
date: "6/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
```

This session:
- filter() to subset data by rows
- *_join() to merge data frames together (analagous to VLOOKUP)

```{r input}
fish <-read_csv(here("data","fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"), sheet="abur") #Only read in sheet called "abur"
```

# dplyr::filter(): conditionally subset by rows

## Filter rows by matching a single character string
- Keep all data observations where fish == "garibaldi"

```{r}
fish_garibaldi <- fish %>%
  filter(common_name == 'garibaldi')
View(fish_garibaldi)
```

```{r}
fish_mohk <- fish %>%
  filter(site == "mohk")
View(fish_mohk)
```

## Filter rows based upon numerical conditions using operators

Observations where total_total count >= 50

```{r}
fish_over_50 <- fish %>%
  filter(total_count >= 50)
View(fish_over_50)
```

Using OR operators (this OR that OR that). Example: garibaldi or blacksmith or black surfperch

Method 1: using | operator (logical OR)

```{r}
fish_3sp <- fish %>%
  filter(common_name == "garibaldi" |
           common_name == "blacksmith" | 
           common_name == "black surfperch")
View(fish_3sp)
```

Method 2: using %in% operator, as long as you are filtering within the same variable. 

```{r}
fish_3sp_v2 <- fish %>%
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
View(fish_3sp_v2)
```

```{r}
fish_gar_2016 <- fish %>%
  filter(year == 2016 | common_name == "garibaldi")
View(fish_gar_2016)
```

## AND operator (i.e., both must be true)

We use , (comma) or & (ampersand) for AND operator

```{r}
aque2018 <- fish %>%
  filter(year == 2018, site == "aque")
View(aque2018)
```
  
The same thing can be done with & instead of ,

```{r}
aque2018_2 <- fish %>%
  filter(year == 2018 & site == "aque")
View(aque2018_2)
```

Or just two filter steps in sequence:

```{r}
aque_2018_3 <- fish %>%
  filter(year ==2018) %>%
  filter(site == "aque")
View(aque_2018_3)
```

```{r}
low_gb_wr <- fish %>%
  filter(common_name %in% c("garibaldi", "rock wrasse") &
                              total_count <= 10))
View(low_gb_wr)
```

### Partial patterns using stringr::str_detect()

Example: keep observations where the common name contains string pattern "black"

```{r}
fish_bl <- fish %>%
  filter(str_detect(common_name, pattern="black"))
View(fish_bl)
```

```{r}
fish_it <- fish %>%
  filter(str_detect(common_name, pattern="it", negate=TRUE)) #negate=TRUE inverts selection
View(fish_it)
```

## dplyr::*_join() to merge data frames

- full_join(): returns all rows and all columns from both x and y
  - "safest" option
- left_join(): return all rows from x, and all columns from x and y
  - rows in x with no match in y will get NA values in new columns
  - if multiple matches between x and y, all combinations of matches are returned
- inner_join(): returns all rows from x where there are matching values in y, and all columns from x and y
  - drops observations w/o match between merged data frames
  
In this module, we will use kelp_abur as "left" data frame and fish as our "right" data frame.

```{r}
abur_kelp_fish <- kelp_abur %>%
  full_join(fish, by =  c("year", "site"))
View(abur_kelp_fish)
```

```{r}
kelp_fish_left <- kelp_abur %>%
  left_join(fish, by = c("year", "site"))
View(kelp_fish_left)
```

```{r}
kelp_fish_injoin <- kelp_abur %>%
  inner_join(fish, by = c("year", "site"))
View(kelp_fish_injoin)
```

```{r}
my_fish_join <- fish %>%
  filter(site == "abur") %>%
  left_join(kelp_abur, by= c("year","site")) %>%
  mutate(fish_per_frond = total_count / total_fronds)
View(my_fish_join)
```

## HTML table with kable() and kableExtra

Can make nice table in HTML with knitr::kable and functions in kableExtra

```{r}
my_fish_join %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width=FALSE)
```
Other options available by Googling kableExtra.
