---
title: "Tidying"
author: "Max Kagan"
date: "6/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages

library(tidyverse)
library(here)
library(janitor)
library(readxl)
```

```{r readin}
inverts <- read_excel(here("data", "inverts.xlsx"))
```

### Going from wider to longer format.

Year variable is spread over three columns (so this is in "wide" format). If we want to get in "long" format (gather together observations spread over multiple columns), we can do this with tidyr::pivot_longer().

We want to take the data from columns "2016", "2017", and "2018", and make into two new columns:
- Year
- sp_count (number of species observed)

```{r}
inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")
```

Class of "year" is actually character, not numeric. To make it numeric, we can use dplyr::mutate.

```{r}
inverts_long <- inverts_long %>%
  mutate(year = as.numeric(year))
```


#### Going from longer to wider format.

Let's say we want each species to be its own column (right now they are in a single column called "common_name". To do this, we use pivots_longer.

```{r}
inverts_wide <- inverts_long %>%
  pivot_wider(names_from = common_name,
              values_from = sp_count)
```

The species names are not in the most coder-friendly format, though, as we can see.

The janitor package is useful for quick data cleaning exercises:
- janitor::clean_names(): update column headers to case of your choosing
  - default is lower_snake_case
  - other options include:
    - "snake"
    - "lower_camel"
    - "upper_camel"
    - screaming_snake"
    - lower_upper
    - upper_lower
- janitor::get_dupes(): see all rows that are duplicates within variables you chose
- janitor::remove_empty(): remove empty rows and/or columns
- janitor:: adorn_*(): jazz up tables

```{r}
inverts_wide <- inverts_wide %>%
  clean_names()
names(inverts_wide)
```

### Separate or uniting items from one/multiple columns

Tidyr::unite() can combine information from multiple columns into a single column. (Basically, like Excel concatenate). It takes the following arguments:
  - data: df
  - col: name of new column
  - columns you are uniting
  - sep: separating symbol/value/character between the two items
  
```{r}
inverts_unite <- inverts_long %>%
  unite(col = "site_year",  # name of new column
        c(site, year), # columns to unite
        sep= "_" # How to separate the columns
  )
View(inverts_unite)
```

```{r}
inverts_moyr <- inverts_long %>%
  unite(col = "mo_yr",
        c(month,year),
        sep="/")
View(inverts_moyr)
```

Separating columns using tidyr::separate()

```{r}
inverts_sep <- inverts_unite %>%
  tidyr::separate(site_year, into=c("my_site", "my_year"))
view(inverts_sep)
```

Replacing a pattern (i.e., find-replace) with stringr::str_replace()

```{r}
ca_abbrv <- inverts %>%
  mutate(
    common_name =
      str_replace(common_name,
                  pattern= "california",
                  replacement = "CA")
  )
View(ca_abbrv)
```