---
title: "Pivot tables with dplyr"
author: "Max Kagan"
date: "6/12/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
```
```{r readin}
lobsters <- read_xlsx(here("data/lobsters.xlsx"), skip=4)
```

Reviewing data with skimr

```{r}

skimr::skim(lobsters)
```

Our goal: report how average size of lobsters has changed for each site, over time.

Key to do this in dplyr is pipe operator %>%

#### Grouping by one variable

1. Group by year
2. Summarize by count (using function n() )

```{r}
lobsters %>%
  group_by(year) %>%
  summarize(count_by_year = n())
```

#### Group by multiple variables

Note: need to remove 'NA's first

```{r}

siteyear_summary <-
  lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear = n(),
            mean_size_mm = mean(size_mm, na.rm=TRUE),
            sd_size_mm = sd(size_mm, na.rm=TRUE)
  )

siteyear_summary
```

kable() can be used for table formatting

```{r}
siteyear_summary %>% 
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report. Using the backtick allows you to write R code inside of Markdown.

```{r}
siteyear_summary <- 
  lobsters %>%
  group_by(site, year) %>%
  summarize(count_by_siteyear = n(),
             mean_size_mm = mean(size_mm, na.rm=TRUE),
            sd_size_mm = sd(size_mm, na.rm=TRUE),
            median_size_mm = median(size_mm, na.rm=TRUE)
)

ggplot(siteyear_summary, 
       aes(x = year, y = median_size_mm, color=site)) + geom_line()
)
```
