---
title: "Creating graphs in R with 'ggplot2'"
author: "Max Kagan"
date: "6/12/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R and it's going to be amazing. 

```{r libraries, echo=FALSE}
library(tidyverse)
library(readxl)
library(here)
```

```{r data1}
ca_np <- read_csv(here("data", "ca_np.csv"))
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

```{r first_plot}
ggplot(data = ci_np, aes(x=year, y=visitors)) +
  geom_line()
```

After we have made the plot an object, we can add / modify it with geom_*:

```{r base_mod1}
gg_base <- ggplot(data = ci_np, aes(x=year, y=visitors))
gg_base + geom_point()
gg_base + geom_col()
gg_base + geom_area()
```

Customizing ggplots:

- color (point / line colors)
- fill (fill color for objects w/ areas)
- linetype (dashed, long dash)
- pch (point style)
- size (element size for points, line thickness)
- alpha (element opacity)

```{r custom1}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

With points, pch (plotting character) can set the type of points.

```{r custom2}
gg_base +
  geom_point(
    color = "purple",
    pch = 17,
    size = 4,
    alpha = 0.5
  )
```

#### Mapping variables unto aesthetics

If updating aesthetics based upon variables, put argument _inside_ of aes.
If outside of aes, will be the same for all.

```{r}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )
```



```{r}
gg_base + 
  geom_col (aes (fill=visitors))
```

#### Built-in themes for ggplot2

- theme_light()
- theme_minimal()
- theme_bw()

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
```


#### Updating axis labels and titles using labs()

Can also update formatting of axis values using scales package.

```{r}
gg_base + 
  geom_line(linetype = "dotted") + 
  theme_bw() + 
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation", 
    subtitle = "1963-2016"
    )
```


#### Combining compatible geoms

Can layer geoms on top of one another to customize

```{r}
gg_base + 
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size=year),
             alpha = 0.5)
```

```{r}
gg_base + 
  geom_col(fill = "orange",
           color = "purple") +
  geom_line(color = "green")
```

#### Multi-series ggplot graphs

How to plot visitation for all California National Parks? We need aesthetic that lets us know how to gruop things, using grouping argument (group = ) or updating aesthetic based upon park_name.

```{r}
gg_np <- ggplot(data=ca_np, aes(x = year, y=visitors, group=park_name))

gg_np +
  geom_line(aes(color=park_name)) + theme_minimal()
```

#### Faceting ggplot graphs

```{r}
gg_np + 
  geom_line(show.legend=FALSE) +
  theme_light() + 
  labs(x = "year", y = "annual visitors") + 
  facet_wrap( ~ park_name)
```

#### Saving gg plot graphs

Graphs can be saved with ggsave()

Arguments:
- width
- height
- dpi

```{r}

ggsave(here("figures", "np_graph.jpg"), dpi = 180, width=8, height = 7)
```
